---
title: "Problem 2.9"
author: "Nicole Higgins"
date: "02/02/2021"
output: html_document
---

#### Enter the matrix A
We have to load the pracma package in order to get `rref`.
```{r, message=FALSE}
require(pracma)
```

Here is the matrix A for this problem.
```{r}
A = cbind(c(12,-7,9,-4,8),c(10,-6,9,-3,7),c(-6,4,-9,-1,-5),c(8,-5,9,0,6),c(4,-7,9,-8,1),c(-18,16,-27,9,-12))
A
```
#### (a) Linear Dependence

Show that the columns of A are linearly independent by finding two different dependency relations among them. You can write your answer a form  like 5 a1+ 4 a2 + 3 a3 + 2 a4 + a5 = 0, where a1, a2, etc are the columns of A. 

```{r}
rref(A)
```
Given the RREF of *A*, we can determine:  
*a1* = -*a4* - *a6*  
and  
*a2* = *a4* + 2*a6*  

#### (b) Ax = b

Augment A with b and show that A x = b is consistent and has infinitely many solutions.

```{r}
Ab = cbind(c(12,-7,9,-4,8),
           c(10,-6,9,-3,7),
           c(-6,4,-9,-1,-5),
           c(8,-5,9,0,6),
           c(4,-7,9,-8,1),
           c(-18,16,-27,9,-12),
           c(14,-12,9,-15,6))
Ab
rref(Ab)
```
The RREF of *Ab* does not reveal any inconsistencies.  
Since *a4* and *a6* are free, there are infinitely many solutions.

#### (c) Remove the redundancies

Remove the free-variable columns from A to get a new, smaller matrix A'. Show that A' x = b has a unique solution and say what that solution is.

```{r}
B = cbind(c(12,-7,9,-4,8),
           c(10,-6,9,-3,7),
           c(-6,4,-9,-1,-5),
           c(4,-7,9,-8,1))
Bb = cbind(c(12,-7,9,-4,8),
           c(10,-6,9,-3,7),
           c(-6,4,-9,-1,-5),
           c(4,-7,9,-8,1),
           c(14,-12,9,-15,6))
B
rref(Bb)
```
Thus, the unique solution to *B* x = *b* is:  
*B1* = 3  
*B2* = -2  
*B3* = 1  
*B4* = 1