---
title: "PSA3"
author: "Nicole Higgins"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pracma)
```

## 3.8

```{r}
StP = c(.85,.09,.06)
Roch = c(.30,.60,.10)
Dul = c(.35,.05,.60) 
M = cbind(StP,Roch,Dul)
```
**a.**  
```{r}
v0 = c(20,20,20)
v1 = M %*% v0
v1
```
M represents the distribution of each cities' rented cars a day later.  
The columns represent the starting city and the rows where they end up, both in the order StP > Roch > Dul.  
Thus, applying M to v0 predicts said distribution a day later as each starting amount is applied to its corresponding column and added. 

**b.**
```{r}
v2 = M %*% v1
v2
M2 = M %*% M
M2 %*% v0
```
**c.**
```{r, warning=FALSE}
dist <- function(today, nxtDay) {
  n <- 0
  while(round(today, 1) != round(nxtDay, 1)) {
    today <- nxtDay
    nxtDay <- M %*% today
    n <- n + 1
  }
  print(round(nxtDay, 1))
  n
}

dist(0, v0)
```
it took v0 9 applications to stabilize

**d.**
```{r, warning=FALSE}
dist(0, c(60,0,0))
dist(0, c(40,10,10))
dist(0, c(5,10,45))
dist(0, c(41,10,9))
```
they do stabilize, the ones closer to the stabilization point taking fewer applications of M to do so.

## 3.9
```{r}
A = rbind(c( 0 , 1 , 0 , 1 , 1 ,0), 
          c(1 , 0 , 1 , 1 , 0, 0 ),
          c( 0 , 1 , 0 , 1 , 1, 0 ),
          c( 1 , 1 , 1 , 0 , 1, 0 ),
          c( 1 , 0 , 1 , 1 , 0, 1 ), 
          c(0, 0, 0, 0, 1, 0))
A
```
**a.**
```{r}
v = c(1,1,1,1,1,1)
A %*% v
```
this vector tells us how many connections each vertex has  
  
**b.**
```{r}
A %*% A
```
**c.**  
the (2,5) entry tells us 2 and 5 share 3 connections  
the (2,3) entry tells us 2 and 3 share 1 connection   
the (2,6) entry tells us 2 and 5 share no connections


